Mosquito parameter fitting
==========================

This script is used to (re)fit a Briere function to published mosquito data to 
generate best estimates and uncertainty in key variables that are dependent 
on temperature, humidity and rainfall. This is done to replace polynomial 
fitting previously used and determine the values of the underlying parameters 
which were not published. 

The key parameters are:
* Biting rate
* Infection probability
* Oviposition rate
* Pupae transition rate
* Transmission probability
* Extrinsic incubation rate

For each of these variables estimates and uncertainty for the underlying 
parameters are produced. 

```{r setup}
# Restart R and set working directory to source file location

library(tidyverse)
library(LeftysRpkg)
library(tidymodels)
library(minpack.lm)

# User inputs - change these for different variables
parameter <- "transmission_prob"

# For specified parameter set file of data and constraints on tmax parameter
# of the Briere function - constraints set based on guess and check and the 
# fits in the paper Mordecai et al (2017) paper (where raw data came from)

if (parameter == "biting_rate") {
  fileName <- "Biting_rate-Extraction.csv"
  tminMin <- 10; tminMax <- 25
  tmaxMin <- 35; tmaxMax <- 50
  mMin <- 0; mMax <- 10
  aMin <- 0; aMax <- 0.000087
} else if (parameter == "infection_probability") {
  fileName <- "infection_probability_data.csv"
  tminMin <- 16; tminMax <- 25
  tmaxMin <- 36; tmaxMax <- 45
  mMin <- 0; mMax <- 10
  aMin <- 0; aMax <- 0.00098
} else if (parameter == "oviposition_rate"){
  fileName <- "Egg_per_female_per_day-Extractions.csv"
  tminMin <- 11; tminMax <- 21
  tmaxMin <- 32; tmaxMax <- 42
  mMin <- 0; mMax <- 10
  aMin <- 0; aMax <- 0.009
} else if (parameter == "pupaeTransition_rate"){
  fileName <- "Tansition_rate_pupae-Extraction.csv"
  # data extracted from Yang et al. 2009.
  tminMin <- 10; tminMax <- 20
  tmaxMin <- 39; tmaxMax <- 42
  mMin <- 0; mMax <- 10
  aMin <- 0; aMax <- 0.0005
} else if (parameter == "transmission_prob"){
  fileName <- "transmission_probability_data.csv"
  tminMin <- 12; tminMax <- 22 
  tmaxMin <- 37; tmaxMax <- 42 
  mMin <- 0; mMax <- 5
  aMin <- 0; aMax <- 0.0015
} else if (parameter == "virusIncubation_rate"){
  fileName <- "Virus_incubation_rate-Extraction.csv"
  tminMin <- 16; tminMax <- 25
  tmaxMin <- 40; tmaxMax <- 46 
  mMin <- 0; mMax <- 10
  aMin <- 0; aMax <- 0.00007
}

# Source Briere function for fitting 
source(file.path(getwd(), "code", "functions.R"), echo = TRUE)

# Directories and files
dataFile <- file.path(getwd(), "data", fileName)
outputFolder <-  file.path(getwd(), "data", "fitted_parameters")

plotRandom <- FALSE

```

```{r Load specified data}
# Load specified mosquito vs weather data
mosData <- read_csv(dataFile, col_names = FALSE) %>%
        rename(temperature = X1,
               estimate = X2) # generic name

# Round data as extracted manually from figures - what we use for fitting
mosDataClean <- mosData %>%
        mutate(temperature = round(temperature, 3),
          estimate = round(estimate, 3)) %>%
        mutate(estimate = ifelse(estimate < 0.01, 0, estimate))
# Add a data point based on Mordecai et al 2017 paper
if (parameter == "transmission_prob") {
mosDataClean <- mosDataClean %>%
  bind_rows(tibble(temperature = 39, estimate = 0))
} else if (parameter == "infection_probability") {
mosDataClean <- mosDataClean %>%
  bind_rows(tibble(temperature = 40, estimate = 0))  
} else if (parameter == "virusIncubation_rate") {
mosDataClean <- mosDataClean %>%
  bind_rows(tibble(temperature = 45, estimate = 0))  
}

```

# Best estimate parameters

First we do a non-linear least squares fit using a Briere function. 

```{r Best estimate fitting }
# Guess and check to fine tune initial conditions
tseq <-  mosDataClean$temperature
qplot(tseq, briere(tseq, 18, 39, 2, 0.00075), 
      geom = "line", 
      ylim = c(0, NA), 
      xlim = c(0, 50), 
      ylab = paste(parameter)) + 
        geom_point(aes(y = mosDataClean$estimate), colour = "blue") +
        PlotOptions()

if (parameter == "biting_rate") {
  initialParList <- list(tmin = 10, tmax = 45, m = 2, a = 0.00041)
} else if (parameter == "infection_probability"){
  initialParList <- list(tmin = 15, tmax = 39, m = 2, a = 0.00098)
} else if (parameter == "oviposition_rate"){
  initialParList <- list(tmin = 10, tmax = 40, m = 3, a = 0.005)
} else if (parameter == "pupaeTransition_rate"){
  initialParList <- list(tmin = 10, tmax = 40, m = 3, a = 0.0005)
} else if (parameter == "virusIncubation_rate"){
  initialParList <- list(tmin = 10, tmax = 45, m = 2, a = 0.000064)
} else if (parameter == "transmission_prob"){
  initialParList <- list(tmin = 18, tmax = 39, m = 2, a = 0.00075)
} 

# Fit model --------------------------------------------------------------------
# First with all parameters being free 
fitModel1 <- nlsLM(estimate ~ briere(temperature, tmin, tmax, m, a), 
                 data = mosDataClean, 
                 start = initialParList,
                 lower = c(tminMin, tmaxMin, mMin, aMin),
                 upper =  c(tminMax, tmaxMax, mMax, aMax),
                 control = nls.control(maxiter = 1000))
summary(fitModel1)

# BriereFunction2 <- function(temp, tmin, m, c) {
#   return(ifelse(t < tmin | t > 40, 0, c*t*(t-tmin)*(40-t)^(1/m)))
#   
# }
# # Second fit force tmin = 13, tmax = 40 for exploration (Not used further)
# fitModel2 <- nlsLM(estimate ~ briere(temperature, tmin, 40, m, a),
#                  data = mosDataClean,
#                  start = initialParList[c(1,3:4)],
#                  control = nls.control(maxiter = 1000))
# summary(fitModel2)

bestModel <- fitModel1 # choose free model

# Plot best fit model with new data --------------------------------------------
fitModelData <- data.frame(temperature = seq(5, 45, length = 100),
  fitvalue = predict(bestModel, list(temperature = seq(5, 45, length = 100))))

ggplot(data = fitModelData, aes(x=temperature, y = fitvalue)) +
        geom_line(colour = "black") +
        geom_point(data = mosDataClean, aes(x = temperature, y = estimate),
                   colour = "blue") +
        labs(x = "Temperature", y = paste(parameter)) +
        PlotOptions()

# Statistics for best fit model ------------------------------------------------
summary(bestModel)
bestEstimates <- coef(bestModel)
standardErrors <- summary(bestModel)$coefficients[, 2]
sd <- standardErrors*sqrt(nrow(mosDataClean))
# Assuming a normal distribution can get 95% confidence intervals
lower95 <- bestEstimates - 1.96*standardErrors
upper95 <- bestEstimates + 1.96*standardErrors

# Store statistics and save 
bestFitStats <- tibble(parameter = names(bestEstimates), 
  estimates = bestEstimates,
  se = standardErrors,
  variance = sd^2,
  lower95 = lower95,
  upper95 = upper95)

```

First we will look at the Briere functions produced when we simply sample from 
the estimated 95% CIs for the fitted parameters. 

```{r Plot with sampled values}
if (plotRandom) {
# Create samples for each parameter using the 95% CI
nsamples <- 100
paramSamples <- tibble(tmin = runif(nsamples, min = lower95["tmin"], 
  max = upper95["tmin"]),
  tmax = runif(nsamples, min = lower95["tmax"], max = upper95["tmax"]),
  a = runif(nsamples, min = lower95["a"], max = upper95["a"]),
  m = runif(nsamples, min = lower95["m"], max = upper95["m"])) 

# a has to be > 0 so reset values of a < 0 to 0
paramSamples <- paramSamples %>%
  mutate(a = ifelse(a < 0, 0, a))

for (ii in 1:nsamples) {
  sample <- paramSamples[ii,]
  newFitValue <- briere(fitModelData$temperature, 
    sample$tmin, sample$tmax, sample$m, sample$a)
  
  fitModelData[[paste0("sample", ii)]] <- newFitValue
  
}

# Plot all the sampled results
ggplot(data = gather(fitModelData, "sample", "value", 2:ncol(fitModelData)), 
  aes(x= temperature, group = sample, y = value)) +
  geom_line() + 
  # geom_point(data = mosDataClean, aes(x = temperature, y = estimate), 
  #   colour = "blue") +
  coord_cartesian(ylim = c(0, 1)) +
  labs(x = "Temperature", y = paste(parameter)) +
  PlotOptions()
}

```

# Bootstrapping

Now try to generate uncertainty using a bootstrap. 

```{r Function for bootstrap fitting}

Fit_on_bootstrap <- function(split) {
  
  modelFitSplit <- tryCatch(nlsLM(estimate ~ briere(temperature, tmin, tmax, m, a), 
    data = analysis(split), 
    start = initialParList,
    lower = c(tminMin, tmaxMin, mMin, aMin),
    upper = c(tmaxMax, tmaxMax, mMax, aMax), 
    control = nls.control(maxiter = 1000, 
      warnOnly = TRUE, scaleOffset = 1,
      minFactor = 1/4096)), 
    error = function(err) {
      print("nlsLM won't fit on boot")
      # Return linear model which will NULL model
      fitData <- analysis(split)
      errModelSplit <- lm(estimate ~ temperature, data = analysis(split))
      return(errModelSplit)})
  
  # Return final model for split
  return(modelFitSplit)
}

```

```{r Bootstrap estimates}

# Set up bootstraps
bootSeed <- sample(1000:3000, 1)
# bootSeed <- 2226
set.seed(bootSeed)

# Will not always get good fits so do enough boots so that numSets can be 
# obtained as a lookup table
numBoots <- 35000 
numSets <- 10000

boots <- bootstraps(mosDataClean, times = numBoots, apparent = TRUE)

# Generate bootstrapped models
boot_models <- boots %>% 
        mutate(model = map(splits, Fit_on_bootstrap),
               coef_info = map(model, tidy),
               predictions = map(model, augment))

params_boot <- boot_models %>% 
        unnest(coef_info) %>% 
        select(id, term, estimate) %>%
        spread(term, estimate) %>% 
        # remove problematic bootstrap fits
        filter(is.na(`(Intercept)`), is.na(temperature), a > aMin, m > mMin, 
              tmax > tmaxMin, tmax < tmaxMax, 
              tmin > tminMin, tmin < tminMax) %>%
        select(-`(Intercept)`, -temperature)

# Print out percentage of bootstraps that had problematic fits for checking
100*(numBoots - nrow(params_boot))/numBoots

# Calculate parameter statistics
params_stats <- params_boot %>% 
        gather("term", "estimate", 2:ncol(.)) %>%
        filter(id != "Apparent") %>% # remove best estimates for stats
        group_by(term) %>%
        summarise(average = mean(estimate),
                  lower = quantile(estimate, 0), 
                  lower95 = quantile(estimate, 0.025),
                  lowerIQR =  quantile(estimate, 0.25), 
                  median = median(estimate),
                  upperIQR =  quantile(estimate, 0.75), 
                  upper95 = quantile(estimate, 0.975),
                  upper = quantile(estimate, 1))

# Select numSets for lookup table - set aside Apparent first and then add as 
# best estimates

params_table <- params_boot %>% 
  filter(id != "Apparent") %>%
  sample_n(numSets, replace = FALSE) %>%
  bind_rows(params_boot %>% filter(id == "Apparent")) %>%
  arrange(id)

# Filter out problematic bootstrap models
boot_models <- boot_models %>% filter(id %in% params_boot$id)

```

```{r Plot bootstrap fits}
# New data for smoothing 
new_preds <- tibble(temperature = seq(5, 45,  length.out = 250))

preds <- augment(bestModel, newdata = new_preds, 
                 type.predict = "response")

# Create smoother predictions for bootstrapped replicate
preds <- boot_models %>%
        mutate(new_predictions = map(model, augment, newdata = new_preds, 
                                     type.predict = "response")) %>%
        unnest(new_predictions) %>%
        # group by each value of xseq and get quantiles
        group_by(., temperature) %>%
        summarise(lwr_CI = quantile(.fitted, 0.025),
                  lwr_IRQ = quantile(.fitted, 0.25),
                  upr_IQR = quantile(.fitted, 0.75),
                  upr_CI = quantile(.fitted, 0.975)) %>%
        ungroup() %>%
        merge(., preds, by = 'temperature')

# Plot model fit and bootstrap ranges
bootstrapPlot <- ggplot() + 
        geom_ribbon(data = preds, aes(x = temperature, ymin = lwr_CI, ymax = upr_CI,
                                      fill = "ci",  alpha = "ci")) +
        geom_ribbon(data = preds, aes(x = temperature, ymin = lwr_IRQ, ymax = upr_IQR, 
                                      fill = "iqr",  alpha = "iqr")) +
        geom_point(data = mosDataClean, aes(temperature, estimate, colour = "data")) +
        geom_line(data = preds, aes(x = temperature, y = .fitted, colour = "fit")) +
        ylab(paste(parameter)) +
        xlab('Temperature') +
        scale_colour_manual(NULL, values = c("blue", "black"),
                            limits = c("data", "fit"),
                            labels = c("Data", "Fitted values")) +
        scale_alpha_manual(NULL, values = c(0.4, 0.1),
                           limits = c("iqr", "ci"),
                           labels = c("Fitted IQR", "Fitted 95% CI")) +
        scale_fill_manual(NULL, values = c("black", "black"),
                          limits = c("iqr", "ci"),
                          labels = c("Fitted IQR", "Fitted 95% CI")) +
        expand_limits(y = 0) +
        PlotOptions() + theme(legend.spacing = unit(0, "cm"),
                              legend.margin = margin(0, 0, 0, 0, "cm")) +
        guides(colour = guide_legend(order = 1),
               alpha = guide_legend(order = 2),
               fill = "none")

bootstrapPlot

```


```{r Save the outputs}
save(bootSeed, bestFitStats, params_boot, params_stats, params_table, boot_models, 
  bootstrapPlot, file = file.path(outputFolder, paste0(parameter, "_fits.Rda")))

write_csv(bestFitStats, file = file.path(outputFolder, paste0(parameter, "_best_fit.csv")))

write_csv(params_table, file = file.path(outputFolder, paste0(parameter, "_sets.csv")))

write_csv(params_stats, file = file.path(outputFolder, paste0(parameter, "_statistics.csv")))
```

```{r}

load(file.path(getwd(), "data", "fitted_parameters" , paste0(parameter, "_fits.Rda")))

```

